---
title: "PARE: PARtial Embeddings for deconfounded dimension reduction"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Maintainer**: Andrew Chen, andrewac@pennmedicine.upenn.edu

**License**: Artistic License 2.0

Partial embeddings (PAREs) provide a framework for removal of confounding effects from any distance-based dimension reduction method including $t$-SNE, UMAP, Laplacian Eigenmaps, diffusion map embeddings, and others. For Euclidean distances, PAREs are equivalent to applying dimension reduction to the principal coordinates after regressing out nuisance covariates.

## 1. Installation
The R package can be installed via devtools by running the following code

```
# install.packages("devtools")
devtools::install_github("andy1764/PARE", build_vignettes = TRUE)
```

Then, you can load this package via

```
library(PARE)
```

## 2. Usage
`PARE` leverages the existing R implementations for dimension reduction methods. Below is an example call for partial UMAP in the `iris` dataset, which removes the species effects:

```{r umap}
library(PARE)
library(umap)

pe <- pare(iris[,-5], ~ Species, data = iris, umap)
```

A vignette is provided for the `pare` function, which also covers other dimension reduction methods

```
vignette("pare")
```

## 3. Citations
If using the PARE methodology, please cite the following paper:

> Chen, A. A., Clark, K., Dewey, B., DuVal, A., Pellegrini, N., Nair, G., Jalkh, Y., Khalil, S., Zurawski, J., Calabresi, P., Reich, D., Bakshi, R., Shou, H., Shinohara, R. T., Initiative,  the A. D. N., & Cooperative,  the N. A. I. in M. S. (2023). Deconfounded Dimension Reduction via Partial Embeddings (p. 2023.01.10.523448). bioRxiv. https://doi.org/10.1101/2023.01.10.523448
